---
title: "Exploratory analysis of exported cases"
author: "Sangeeta Bhatia and Gina Cuomo-Dannenburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    fig_width: 7
    fig_height: 5
    fig_caption: true
    highlight: "tango"
    reference_docx: word_style.docx
---

```{r exported-cases-eda-1 }

## orderly::orderly_run("exported_cases_eda")

## Read directly from google drive to avoid having to download the
## data. Note that this step will prompt for google sign in and
## ask you to allow Tidyverse API to access your srive.
googlesheets4::sheets_deauth()
exported_raw <- googlesheets4::read_sheet(
    "https://docs.google.com/spreadsheets/d/1xXv_TT5OarKhl_bcQJE9fYizHYjNViUlG7y5f1c0SEU/edit#gid=207824382",
    sheet = "by_country",
    col_types = "c", ## read everything as character, easier than having to guess excel's date format
    na = c("", "NA") ## sometimes people actually write NA rather than leaving a cell blank.
    )

exported <- janitor::clean_names(exported_raw)

```

## Make dates

```{r exported-cases-eda-2}

exported$date_report_dd_mm_yyyy <-
    lubridate::dmy(exported$date_report_dd_mm_yyyy)

    
```

```{r exported-cases-eda-3 }
exported$date_travelled_from_wuhan_dd_mm_yyyy <-
    lubridate::dmy(exported$date_travelled_from_wuhan_dd_mm_yyyy)
    
```

```{r exported-cases-eda-4 }
## Two of these are unclear -  a) > 20/01/2020, b) after 22nd
exported$date_onset_symptoms_dd_mm_yyyy <-
    lubridate::dmy(exported$date_onset_symptoms_dd_mm_yyyy)

```

```{r exported-cases-eda-5 }
## As of 02022020. Again, 2 are unclear.
exported$date_hospitalized_dd_mm_yyyy <-
    lubridate::dmy(exported$date_hospitalized_dd_mm_yyyy)

```

```{r exported-cases-eda-6 }

exported$date_confirmed_dd_mm_yyyy <-
    lubridate::dmy(exported$date_confirmed_dd_mm_yyyy)

```

```{r exported-cases-eda-7 }

exported$date_death_dd_mm_yyyy <-
    lubridate::dmy(exported$date_death_dd_mm_yyyy)

```

```{r exported-cases-eda-8 }

exported$date_recovered_dd_mm_yyyy <-
    lubridate::dmy(exported$date_recovered_dd_mm_yyyy)

```

Additional columns - date of outcome (death or discharge). NA if
unknown. 


Check that there are no cases for which dates are not NA.

```{r exported-cases-eda-9 }
## this should be 0 rows.
exported[!is.na(exported$date_recovered_dd_mm_yyyy) &
         !is.na(exported$date_death_dd_mm_yyyy), ]

```

## Coding outcome

```{r exported-cases-eda-10 }
## outcome - if death of recovery is available, discharged
## if date of death is available, death
## if none of the two dates are available, other.
exported$outcome <- dplyr::case_when(
    
    !is.na(exported$date_recovered_dd_mm_yyyy) ~ "recovery",
    !is.na(exported$date_death_dd_mm_yyyy) ~ "death",
    TRUE ~ "other"
    
)

exported$date_outcome <- dplyr::case_when(
    
    exported$outcome == "recovery" ~ exported$date_recovered_dd_mm_yyyy,
    exported$outcome == "death" ~ exported$date_death_dd_mm_yyyy,
    exported$outcome == "other" ~ Sys.Date()
    
 )

exported$outcome_censored <- dplyr::case_when(

    exported$outcome == "recovery" ~ 0,
    exported$outcome == "death" ~ 0,
    TRUE ~ 1
    
    )

```

## Fix age

```{r exported-cases-eda-11 }
## 20s --> 25
## 30s --> 35 and so on.
idx_with_s <- grep("s", exported$age_years)
## Remove s and add 5.
exported$age_years <- gsub(
    pattern = "s", replacement = "", x = exported$age_years
)
## > 50, 50+,
idx_with_signs <- grep(">", exported$age_years, fixed = TRUE)
idx_with_signs <- c(
    idx_with_signs,
    grep("+", exported$age_years, fixed = TRUE)
 )

exported$age_years <- gsub(
    pattern = ">", replacement = "", x = exported$age_years
)

exported$age_years <- gsub(
    pattern = "+", replacement = "", x = exported$age_years, fixed = TRUE
)

## Make integer and don't worry about rows that don't parse.
exported$age_years <- as.integer(exported$age_years)

## Add 5 where we have changed stuff.
exported$age_years[idx_with_signs] <- exported$age_years[idx_with_signs] +
    5

exported$age_years[idx_with_s] <- exported$age_years[idx_with_s] +
    5


```

```{r exported-cases-eda-12 }

both_dates_available <- dplyr::filter(exported, !is.na(date_travelled_from_wuhan_dd_mm_yyyy) &
                                                !is.na(date_onset_symptoms_dd_mm_yyyy))

## 03022020. 86 have both dates available.

both_dates_available$onset_after_travel <- dplyr::case_when(
                                          
    both_dates_available$date_travelled_from_wuhan_dd_mm_yyyy >= both_dates_available$date_onset_symptoms_dd_mm_yyyy ~ "travel_on_or_after_onset",
    both_dates_available$date_travelled_from_wuhan_dd_mm_yyyy < both_dates_available$date_onset_symptoms_dd_mm_yyyy ~ "travel_before_onset"
)

```

## How many of these have pneumonia

How many have either yes/no rather than missing data?

```{r exported-cases-eda-13 }
sum(! is.na(both_dates_available$pneumonia_y_n))
```

Group by onset before/after travel

```{r exported-cases-eda-14 }

dplyr::count(both_dates_available, onset_after_travel, pneumonia_y_n)


```
Similarly for case severity classification.

```{r exported-cases-eda-15 }

dplyr::count(both_dates_available, onset_after_travel, classification_mild_moderate_severe)

```

And other symptoms

Fever

```{r exported-cases-eda-16 }

dplyr::count(both_dates_available, onset_after_travel, febrile_y_n)

```

Cough


```{r exported-cases-eda-17 }

dplyr::count(both_dates_available, onset_after_travel, cough_y_n)

```


# country summary sheet
Read in the country summary information so we can output this - Lucy and Ilaria will be using this in their tasks

```{r country-summary-1 }

## orderly::orderly_run("exported_cases_eda")

## Read directly from google drive to avoid having to download the
## data. Note that this step will prompt for google sign in and
## ask you to allow Tidyverse API to access your srive.
googlesheets4::sheets_deauth()
summary_raw <- googlesheets4::read_sheet(
    "https://docs.google.com/spreadsheets/d/1xXv_TT5OarKhl_bcQJE9fYizHYjNViUlG7y5f1c0SEU/edit#gid=207824382",
    sheet = "summary_by_country",
    col_types = "c", ## read everything as character, easier than having to guess excel's date format
    na = c("", "NA") ## sometimes people actually write NA rather than leaving a cell blank.
    )

drops <- c("updated_at")

summary_by_country <- summary_raw[ , !(names(summary_raw) %in% drops)] %>% 
  janitor::clean_names()

```

```{r country-summary-2}

summary_by_country <- summary_by_country %>%
  filter(!is.na(country))

print(summary_by_country$country)

```



Make the time stamp and object and then remove these columns from the cleaned version
```{r country-summary-3}

time_of_update <- summary_raw$updated_at[1]

time_of_update <- lubridate::dmy_hms(time_of_update, tz = "GMT")
print(time_of_update)

```



Numeric columns
```{r country-summary-4}

numeric_col_indices <- grep(c("country|url|url2|notes"), names(summary_by_country), invert = TRUE)

```

Convert to numerics

```{r country-summary-5}
summary_by_country <- summary_by_country %>% 
  mutate_at(names(summary_by_country)[numeric_col_indices], as.numeric)


```


Fix this so that the imputed information cannot be negative -- correct this so that this becomes zero. 
```{r country-summary-6}
## if the number of cases to impute is negative, correct this to be zero

summary_by_country$unassigned_recoveries[summary_by_country$unassigned_recoveries < 0] <- 0


```























<!-- Similar analysis for deaths. -->

<!-- ```{r exported-cases-eda-12 } -->

<!-- deaths <- readr::read_csv("data/deaths.csv") -->
<!-- deaths_long <- tidyr::gather( -->
<!--     deaths, -->
<!--     key = symptom, -->
<!--     value = yes, -->
<!--     pneumonia:underlying_medical_conditions, -->
<!--     factor_key = TRUE -->
<!--  ) -->

<!-- ``` -->

<!-- Most common symptoms among those who have died. -->

<!-- ```{r exported-cases-eda-13 } -->

<!-- dplyr::arrange(na.omit(dplyr::count(deaths_long, symptom, yes)), desc(n)) -->

<!-- ``` -->


<!-- How much time between onset and death (in cases where we know both -->
<!-- dates of course)? -->

<!-- ```{r exported-cases-eda-14 } -->
<!-- out <- dplyr::filter(deaths, ! is.na(date_onset) & ! is.na(date_death)) -->

<!-- sort(lubridate::dmy(out$date_death) - lubridate::dmy(out$date_onset)) -->

<!-- ``` -->
<!-- Similalry from hospitalisation to onset. -->

<!-- ```{r exported-cases-eda-15 } -->
<!-- out <- dplyr::filter(deaths, ! is.na(date_onset) & ! is.na(date_hospitalised)) -->

<!-- sort(lubridate::dmy(out$date_hospitalised) - lubridate::dmy(out$date_onset)) -->

<!-- ``` -->

<!-- And death to hospitalisation. -->

<!-- ```{r exported-cases-eda-16 } -->

<!-- out <- dplyr::filter(deaths, ! is.na(date_death) & ! is.na(date_hospitalised)) -->

<!-- sort(lubridate::dmy(out$date_death) - lubridate::dmy(out$date_hospitalised)) -->

<!-- ``` -->

```{r write-csv-files }

readr::write_csv(x = exported_raw, path = "exported_cases_raw.csv")
readr::write_csv(x = exported, path = "exported_cases_cleaned.csv")
readr::write_csv(x = summary_raw, path = "summary_by_country_raw.csv")
readr::write_csv(x = summary_by_country, path = "summary_by_country_cleaned.csv")
saveRDS(time_of_update, file = "time_exported_case_update.rds")

```
